---
title: "Analyzing Guinea Pig Tooth Growth Data"
author: "Alexander Pyle, <apyle@github.com>"
date: "December 27, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

* Load the ToothGrowth data and perform some basic exploratory data analyses 
* Provide a basic summary of the data.
* Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
* State your conclusions and the assumptions needed for your conclusions. 

Some criteria that you will be evaluated on

* Did you perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
* Did the student perform some relevant confidence intervals and/or tests?
* Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
* Did the student describe the assumptions needed for their conclusions?


## Overview

```{r overview, echo=FALSE}

# load required libraries
library(data.table)
library(ggplot2)
library(xtable)

# variables used in the text and code, reinitiallized in the next chunk. Set here for use in text

```

Does delivery method and/or dosage of vitamin C affect tooth growth in guinea pigs? 
In this paper we will reviews R's 'ToothGrowth' dataset and performs the appropriate 
analysis to answer this question and identify the method and dosage, if any, that 
maximizes tooth growth.

## Data Summary

```{r datasummary1, fig.height=6, fig.width=8, fig.align='center'}

# load the Guinea Pig Tooth Growth data
data(ToothGrowth)

```

The 'ToothGrowth' dataset included in R is the results of giving 60 guinea pigs 
dosages of vitamin C and measuring the resultant tooth growth. The dosages were 
0.5, 1.0, and 2.0 milligrams per day and was delivered either via orange juice 
or asorbic acid. The guinea pigs were separated into six groups of ten and each 
group was given the vitamin C in the same dosage and delivery method. The results 
are summarized below.

```{r datasummary2, fig.height=6, fig.width=8, fig.align='center'}

```


## Analysis

```{r analysis}

```


## Conclusion

```{r conclusion}

```


## Appendix

A summary of the tooth growth data can be found at https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/ToothGrowth.html

This analysis was run with the following configuration.

```{r warning=FALSE}
library(devtools)
devtools::session_info() # display environment the script was create and run in.

```

