---
title: "Analyzing Guinea Pig Tooth Growth Data"
author: "Alexander Pyle, <apyle@github.com>"
date: "December 27, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

* Load the ToothGrowth data and perform some basic exploratory data analyses 
* Provide a basic summary of the data.
* Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
* State your conclusions and the assumptions needed for your conclusions. 

Some criteria that you will be evaluated on

* Did you perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
* Did the student perform some relevant confidence intervals and/or tests?
* Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
* Did the student describe the assumptions needed for their conclusions?


## Overview

```{r overview, echo=FALSE}

# load required libraries
library(data.table)
library(ggplot2)
library(xtable)

# variables used in the text and code, reinitiallized in the next chunk. Set here for use in text

```


## Data Summary

```{r datasummary1, fig.height=6, fig.width=8, fig.align='center'}

# load the Guinea Pig Tooth Growth data
data(ToothGrowth)

```

```{r datasummary2, fig.height=6, fig.width=8, fig.align='center'}

```


## Analysis

```{r analysis}

```


## Conclusion

```{r conclusion}

```


## Appendix

This analysis was run with the following configuration.

```{r warning=FALSE}
library(devtools)
devtools::session_info() # display environment the script was create and run in.

```

